---
title: "EDS-221_Day3_ConditionsAndLoops"
author: "OP"
format: html
editor: visual
---

## Conditional statements

```{r}
temperature <- 95

if (temperature > 97) {
  print("Temperature warning")
}else 
  print("Temperature is within the norm")
```

#### Create an if statement for blue whales lengths,

length = 40 m. Add an if else statement that returns "GIANT" if the length \> 35, "MEDIUM", between 20 and 35, "SMALL" if \< = 20

```{r}

whale_length_m <- 35

if (whale_length_m > 35) {
  print("GIANT")
}else if(whale_length_m > 20 & whale_length <= 35)
  {print("MEDIUM")
}else if (whale_length_m <= 20)
  print("SMALL")

```

### Switch statements

```{r}
housing_type <- "whole"
switch(housing_type, 
       "single" = print("400"),
       "double" = print("750"),
       "whole" = print("1100"))

# the same code in if-else statements
if (housing_type == "single") {
  print("400")
}else if(housing_type == "double"){
  print("750")
}else if (housing_type == "whole"){
  print("1100")
}

```

### For loops

```{r}
height <- c(58, 61, 65, 72, 89, 94, 101)

for (i in seq_along(height)){
  half_height = height[i] / 2
  #print(half_height)
}
  
# --- for each value in the height value, return the sum of the value + the one right after if...

# create blank vector to store the new vector
height2 <- vector (mode = "numeric", length = (length(height)))

for (i in seq_along(height)){
  sum_height = height[i] + height[i+1] 
  height2[i] <- sum_height
}

height2

```

```{r}


for (i in seq_along(height)){
  sum_height = height[i] + height[i+1] 
  print(sum_height)
}

```

```{r}
for (i in seq(1:(length(height) -1))) {
  sum_height = height[i] + height[i+1] 
  print(sum_height)
}

```

#### Create for loop that calculates the value of height for each element divided by the mean of the *entire height vector*

```{r}

for (i in seq_along(height)){
  standardized_height = height[i] / mean(height)
  print(standardized_height)
}
```
